---
title: "Lab 1"
author: "Brad Anderson"
date: "January 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r data and packages}

library(tidyverse)

hp_aggression <- read_csv("H:/Week 1/hp_aggression.csv")

```

##Step 3. Data Wrangling and the Pipe Operator

-select(): select specified columns from a df
-filter(): sekect riws by setting conditions to match
-mutate(): adds columns based on calculations for existing
-arrange(): sorting variables
-group_by(): assign groupings to variable entries, and can perform operations by group
-distinct(): will only retain distinct observations
-head(): select any number of top observations

```{r data_wrangling}

hp_df1 <- hp_aggression %>% 
  select(Name, aggressions, mentions) %>% 
  mutate(rate = aggressions/mentions) %>% 
  group_by(Name) %>% 
  mutate(average_rate = mean(rate)) %>%  #It will caluclate mean rate based on the groups.
  ungroup() %>% 
  select(Name, average_rate) %>% 
  distinct(Name, average_rate) %>% 
  arrange(-average_rate) %>% 
  head(10)
```

##Step 4. ggplot refresher

1. Tell R that you want to use ggplots
2. Tell it what data you want to use (what your x and y are)
3. tell it what type of graphic you want to create

```{r set_levels}

hp_df1$Name <- factor(hp_df1$Name, levels = hp_df1$Name[order(hp_df1$average_rate)]) #This allows you to create ordered levels

```

```{r graph}

ggplot(data = hp_df1, aes(x = Name, y = average_rate))+
  geom_bar(stat = "identity", aes(fill = Name)) + #Quotations are usually used when describing characteristics 
  coord_flip()
```

