---
title: "244 Assignment 1"
author: "Brad Anderson"
date: "February 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


###Task 3: Blues Styles
```{r}
library(tidyverse)
library(nnet)

BluesStyle <- read_csv("H:/244/Assignment 1/BluesStyle.csv")
```

```{r}
BluesStyle$Region<-as.factor(BluesStyle$Region)
BluesStyle$Style<-as.factor(BluesStyle$Style)

```

```{r}
BluesMLR <-multinom(Style ~ Region + BirthYear, data=BluesStyle)
exp(coef(BluesMLR))
```

```{r}

BluesFrame<-data.frame(Region=rep(c("Texas","Delta","East"), 
                              each= 67),
                  BirthYear=rep(seq(1874,1940),3))
```

```{r}

Predprob<-data.frame(BluesFrame, predict(BluesMLR,newdata=BluesFrame, type="probs"))

```

```{r}

prob_gather <-Predprob %>% 
  gather(ThumbStyle, Probability, Alternating:Utility)

```

```{r}

ggplot(prob_gather, aes(x=BirthYear, y=Probability))+
  geom_line(aes(color=ThumbStyle), size=1)+
  facet_wrap(~Region)+
  ggtitle("The Thumb Styles of Blues Guitarists")+
  xlab("Birth Year") +
  ylab("Probability of Using Thumb Style") +
  theme_linedraw()+
  theme(plot.title = element_text(hjust = 0.5))+
  theme(panel.grid.major = element_line(colour = "gray"))+
  theme(panel.grid.minor = element_line(colour = "gray"))+
  guides(color=guide_legend(title="Thumb Picking Style")) +
  ggsave("blues.jpg", width = 6, height = 4, units = c("in"), dpi=300)

```

The	relative	risk	(or	odds)	of	using a Dead thumb style	as	you	switch	
from	the Delta	to	Texas	is	multiplied	by	3.89	(there	is	a	
GREATER	PROBABILITY	of	using	the Dead Thumb Style	if	they are from Texas	versus	
the Delta)

If	birth year	increases	by	1,	the	relative	risk	of	using a Dead thumb style	
versus	attending	using the alternating or utility styles	are	both reduced, meaning as time went on the dead thumb style became less popular in every region.

###Task 4: Mexican Food

```{r}

agaves_v<-rep(1:5, c(30,22,30,109,326))
lilly_v<-rep(1:5, c(6,13,19,83,199))


#both_data<-data.frame(agaves_vector, lilly_vector)
#colnames(both_data)<-c("Los_Agaves", "Lillys")

#agaves_v<-as.numeric(agaves_v)
#lilly_v<-as.numeric(lilly_v)

```

```{r}

f_test <- var.test(agaves_v, lilly_v)

f_test

```

```{r}

tacos_mwu <- wilcox.test(agaves_v, lilly_v)
tacos_mwu

```

There is not a significant difference in median rankings for Lilys Tacos (n=???) and Los Agaves (n = ???), w(65) = 0.49, apha = 0.05.