---
title: "Assignment 3 Lizard"
author: "Brad Anderson"
date: "March 7, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(tidyverse)

#Load the cleaned .csv file.
lizard <- read_csv("lizard.csv")
```

```{r}

#Replace all the "." with NA, which R can handle better.

lizard[lizard =="."]<-NA

#Filter for adult lizards only.
m <- lizard %>% 
  filter(sex == "M" & site == "CALI")
f <- lizard %>% 
  filter(sex == "F" & site == "CALI")

m$weight <- as.numeric(m$weight)
f$weight <- as.numeric(f$weight)

```

```{r}



par(mfrow = c (2,2))
hist(m$weight)
hist(f$weight)
qqnorm(m$weight)
qqnorm(f$weight)

```

```{r}

f_test <- var.test(m$weight,f$weight)
f_test

```
Ther variances are equal, thus I will use student's t-test.

```{r}

t.test_diff <- t.test(m$weight, f$weight, var.equal = TRUE)
t.test_diff

```
Live-trapping of lizards (sampling-with-replacement) in the Northern Chihuahua Desert near White Sands, NM from 1989 to 1990 found that though mean weight for female lizards (n = 79)  was 0.86 grams greater than males (n = 58), the result was not significant (t(130) = -0.7, p = 0.427).

```{r}

all_tails <- lizard %>% 
  filter(site == "CALI") %>% 
  select(sex, tail)

m_broken <- all_tails %>% 
  filter(sex == "M") %>% 
  filter(tail == "B")
#10

m_whole <- all_tails %>% 
  filter(sex == "M") %>% 
  filter(tail == "W")
#46

f_broken <- all_tails %>% 
  filter(sex == "F") %>% 
  filter(tail == "B")
#18

f_whole <- all_tails %>% 
  filter(sex == "F") %>% 
  filter(tail == "W")
#59

male <- c(10, 46)
female <- c(18, 59)

lizard_tails <- rbind(male, female)
colnames(lizard_tails) <- c("Broken" , "Whole")

```

```{r}

prop <- prop.table(lizard_tails, margin = 1)

prop_rounded <- round(prop, 2)

prop_table <- prop.table(prop, margin = 1)
prop_table

```

```{r}
 lizard_chi <- chisq.test(lizard_tails)
lizard_chi
```
There is not a significant difference in the proportion of adult male (n = 56) and female (n = 77) lizards with broken tails ($\chi$^2^(1, n = 133) = 0.31, p = 0.579).
