---
title: "Lab 3"
author: "Brad Anderson"
date: "February 1, 2018"
output: html_document
---
####Bootstrapping

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages and data}

library(tidyverse)
library(boot)
library(nlstools)

penguins <- read_csv("H:/Week 3 Bootstrapping NOLS/penguins.csv")
CellGrowth <- read_csv("H:/Week 3 Bootstrapping NOLS/CellGrowth.csv")

```

Make a vector of values that we will bootstrap.
```{r wrangling}

flip <- penguins %>%
  filter(sex == "FEMALE") %>%
  select(flipper_length)

flip_v <- flip$flipper_length

hist(flip_v)

```

C. Create a function to calculate the means of the bootstrap samples

```{r}

mean_fun <- function (x, i) {mean(x[i])} #i are the different bootstrap samples

```

d. bootstrap flip_v

```{r}

boot_10 <- boot(flip_v, mean_fun, R = 10)
boot_10$t
hist(boot_10$t)
```

```{r}
boot_10000 <- boot(flip_v, mean_fun, R = 10000)
hist(boot_10000$t)
boot_10000
```

e. Find the confidence interval based on the sampling distribution using function boot.ci()

```{r}

penguin_ci <- boot.ci(boot_10000, conf = 0.95, type = "perc")
penguin_ci

```

###Non-Linear Least Squares

b. Look at data
```{r}

ggplot(CellGrowth, aes(x = Time, y = CellCount)) +
  geom_point()
```

C(t) = A/1+Be^=rt  
B= K-C0/C0

Estimate for carrying capacity (k): ~3700
Estimate for C0: ~2000 
Estimate for B: (3700 - 2000)/2000 : ~0.85

Create a linear model between the ln of CellCount and Time for the first 5 observations (these look like they might follow exponential growth). The Slope of that line is an estimate for the growth rate constant.

```{r}

r_est <- lm(log(CellGrowth$CellCount [1:5]) ~CellGrowth$Time[1:5]) #only using observations 1-5
r_est
#From this we have a guess at r: 0.1035
```
Estimate for A (k): ~3700
Estimate for C0: ~2000 
Estimate for B: ~0.85
Estimate for r: 0.1035

NLS for estimating parameter values for logistic growth, starting with our estimates for A, B and r.

C = A/(1+Be^-rt)

```{r}

cell_fit <- nls(CellCount ~ A/(1+B*exp(-r*Time)), 
                data = CellGrowth,
                start = list(A = 3700, B = 0.85, r = 0.1035), #starting values
                trace = TRUE)

```
The left column shows the sum of squares values if the values on the right are used. It converges at the bottom down to the thousandeths. 

Now I will tell r what these paremeters we found are.
```{r}
A <- coef(cell_fit)[1]
B <- coef(cell_fit)[2]
r <- coef(cell_fit)[3]
```

The ony variable we don't have now is time, so we will make a time sequence (from 0 - 20 hours, and I want there to be 100 values within that). Then I'm going to plug that into my model to find the predicted cell population size.

```{r}

time_seq <- seq(0, 20, length =100) #this will give me a range of values from 0 to 20 with 100 values evenly spaced

cell_pred <- A/(1 + B*exp(-r*time_seq))

pred_df <- data.frame(time_seq, cell_pred)
```

Create a graph of the original data + predictive model line

```{r}

ggplot(CellGrowth, aes( x = Time, y = CellCount)) +
  geom_point( color = "blue", size = 3) +
  theme_bw() +
  geom_line(data = pred_df, aes(x = time_seq, y = cell_pred))
```

