---
title: "Lab 7 Variograms and Kriging"
author: "Brad Anderson"
date: "March 1, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library(sf)
library(sp)
library(gstat)
library(rgdal)
library(tidyverse)
library(devtools)
library(tmap)
```

```{r}

lat_lon <- read_csv("lat_lon.csv")
fire_pm25 <- read_csv("fire_pm25.csv")
counties <- st_read(dsn = ".", layer = "california_county_shape_file")

```

3. Combine PM2.5 attribute data with the lat/long information. You frequently must create/combine the data so that R knows it is spatial.

```{r}

 df <- full_join(lat_lon, fire_pm25) #returns all, even if there is no match

df <- df %>% 
  filter(name !="NA") #Gets ride of random description rows at the bottom. filter() is listwise deletion that gets rid of the entire row.

```

4. Convert to simple features spatial data, and we're going to set the CRS
```{r}
df_sf <- st_as_sf(df, coords = c("latitude", "longitude")) #Get data from df, then tell it which columns refer to the coordinates.
                  
#Now the latitude/longitude information is stored in a new geometry column.

st_crs(df_sf) <- 4326 #Sets coordinate reference system.
st_crs(counties) <- 4326
```

5. We want to look at the spatial data at a certain date after the fire. So we are going to filter the dat to only retain air quality observations from 12/10/17.

```{r}
#The geometry data is 'sticky,' meaning it will stay with the other data and we can kinda forget about it.

pm_12_10 <- df_sf %>% 
  filter(summary_date == "12/10/17") %>% 
  select(pm25_davg)
```

You could just use the plot() function
```{r}
plot(pm_12_10)
plot(counties)

```

6. Make a point plot using ggplot with geom_sf
```{r}

ggplot(pm_12_10)+
  geom_sf(data = counties)+
  geom_sf(aes(color = pm25_davg))

```

Have R make a variogram and do some kriging so that I can predict concentrations for the entire state. First, we need to make a variogram that we can draw the kriging from.

7. Create a simplified data fram, set coordinates, and CRS

```{r}

df2 <- df %>% 
  filter(summary_date == "12/10/17") %>% 
  select(longitude, latitude, pm25_davg)

coordinates(df2) <- ~ latitude + longitude
proj4string(df2) <- CRS("+init=epsg:4326")
```

```{r}
 bubble(df2, "pm25_davg")
```

8. Create a variogram, and fit a variogram model.
```{r}

pm_vg <- variogram(pm25_davg ~ 1, df2) # ~1 is normal kriging
plot(pm_vg)
```
Estimates: Nugget ~ 90, Range ~ 300, Sill ~ 450

```{r}

vg_fit <- fit.variogram(pm_vg, vgm(c("Exp", "Sph", "Gau")))
vg_fit

```
It decided the Gau model is the best.

```{r}
plot(pm_vg, vg_fit)
```

9. New grid of pixels is over which we want to interpolate.

```{r}

Long <- seq(-125, -114, length = 100) #These lat/long bound California.
Lat <- seq(32, 42, length = 100)

Grid <- expand.grid(Long, Lat) #Stored now as a dataframe instead of two seperate vectors.
colnames(Grid) <- c("Long", "Lat")

coordinates(Grid) <- ~ Long + Lat #This tells R the dataframe is spatial data.
gridded(Grid) = TRUE

proj4string(Grid) <- CRS("+init=epsg:4326")

plot(Grid)
```
Now we have a grid of areas that we can make predictions over.

10. Kriging
```{r}

pm_krige <- krige(pm25_davg ~1, df2, newdata = Grid, model = vg_fit)

```

11. Visualize Results
```{r}

krige_df <- data.frame(Grid$Long, Grid$Lat, pm_krige$var1.pred)
colnames(krige_df) <- c("Long", "Lat", "Predicted")

krige_gg <- ggplot() +
  geom_tile(data = krige_df, aes(x = Long, y = Lat, fill = Predicted))+
              geom_sf(data = counties, fill = "NA", color = "white")
            
            krige_gg

```

