---
title: "Lab 8 Spatial Point Pattern Analysis & PCA"
author: "Brad Anderson"
date: "March 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}

library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(spatstat)
library(maptools)
```



###Column Graph of Texas Oil Spills

```{r}

 oil_spills <- read_csv("oil_spills.csv")

df <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` < 2017) %>%  #Use the '' quotations to add in a item with a space in the name. 
  group_by(`Accident Year`) %>% #These headings are not data science friendly, it would be better to have a metadata file that has a description based on a shorter header.
  summarise(Loss = sum (`Net Loss (Barrels)`))

colnames(df) <- c("Year", "Loss")

ggplot(df, aes(x = Year, y = Loss)) + 
  geom_col()


```


###Leaflet plot of spilll locations in TX in 2016
```{r}

df_loc <- oil_spills %>% 
  filter(`Accident State` == "TX" & `Accident Year` == 2016) %>% 
  select(Latitude, Longitude, `Net Loss (Barrels)`)

colnames(df_loc) <- c("latitude", "longitude", "net_loss")

```

R does not yet recognize that this is spatial information. We can convert it into simple features data:

```{r}

oil_sf <- st_as_sf(df_loc, coords = c("longitude", "latitude"), crs = 4326)
# Now the spatial information will be retained, even as we filter(), select(), etc. the data.

leaflet(oil_sf) %>% #Leaflet actually automatically will use lat/long data if it can find it in the data.
  addTiles %>% 
  addMarkers
```

 ##tmap plot with Texas state shapefile
 
```{r}

states <- st_read(dsn = ".", layer = "states") # "." because the file is in the folder?

tex_border <- states %>% 
  filter(STATE_NAME == "Texas") %>% 
  #Transforming here so that when we start making polygons and points they align correctly. This data already has a crs assigned, so we must use transform to change it.
  st_transform(4326)

plot(tex_border)

tm_shape(tex_border) +
  tm_polygons()+
  tm_shape(oil_sf)+
  tm_dots(size = 0.3)
```
 
 Does this point pattern follow CSR?
 
 ###Convert the data to spatial points patterns (combination of point data and the bounding window)
 
```{r}
#Convert from spatial features back into a dataframe.
spill_sp <- as(oil_sf, "Spatial") #R still cannot recognize this as data to do point pattern analysis with.
spill_ppp <- as(spill_sp, "ppp") #ppp is a specific type of data that R understands

tx_sp <- as(tex_border, "Spatial") 
tx_owin <- as(tx_sp, "owin") #This is the outer window that I am going to use in my analysis.

all_ppp <- ppp(spill_ppp$x, spill_ppp$y, window = tx_owin) #x and y are the lat/long and the window is the area we are looking at
```
 
We have just taken spatial information and put it into a format that R understands.

###A kernal density plot for overlapping areas around points:
```{r}

 plot(density(all_ppp, sigma = 0.4)) 

```
Spatial randomness is different from spatial eveness.

###Quadrat test for spatial evenness

Are oil spills evenly distributed throughout the state?

```{r}

oil_qt <- quadrat.test(all_ppp, nx = 5, ny = 5) #A total of 5 regions that exist horizontally, 5 vertically
oil_qt
```
A very small p value, thus NOT CSR and NOT evenly distributed.

If you look at the oil spill data for Texas, it doesn't look like CSR data, it looks like clustering in some areas and blank areas in others. 

```{r}
plot(all_ppp)
plot(oil_qt, add = TRUE, cex = 0.4)
```

For a whole square the expected number of events is 14.5, for partial squares it displays the proportional number of events. 

###G-Function for nearest neighbor analysis
```{r}
 r <- seq(0,1, by = 0.01)

#Simulate CSR data

oil_gfun <- envelope(all_ppp, fun = Gest, r =r, nsim=100)
ggplot(oil_gfun, aes(x = r, y = obs)) +
  geom_line(color = "black") +
  geom_line(aes(x = r, y = theo), color = "red")

```
 Our observed data has a higher prop of point pairs with nearest neighbors at shorter distances compared to csr data.

###Nearest neighbor using the L-function (Ripley's K, standardized)
```{r}

r2 <- seq(0,3, by = 0.5)

oil_lfun <- envelope(all_ppp, fun = Lest, r = r2, nsim = 20, global = TRUE) #Instead of just finding a single nearest neighbor, it makes increasing bubbles around each event, thus much more computationally instensive.

ggplot(oil_lfun, aes(x = r2, y = obs, ))+
  geom_line( color = "black")+
  geom_line(aes(x = r2, y = theo), color = "blue")
```
Both L and G functions show that nearest neighbor and global neighbors show that the distance is small, indicating clustering. 
