---
title: "Lab Week 5 Missing Data & Mult Imputation"
author: "Brad Anderson"
date: "February 15, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(tidyverse)
library(VIM) #package for exploring and visualizing missing data
library(mice)

```

```{r data}

df <- read_csv("~/github/244/Lab 5 Missing Data/data/df_missings.csv")

```

##Step 3. 

a. Identify complete cases and missing cases

```{r}

cc <- df[complete.cases(df),] #complete cases with no missing observations
missings <- df[!complete.cases(df),]
```

b. Visually identify the structure of the missingness.

```{r}

aggr(df, prop = F, numbers = T) #prop = T will use proportions instead of numbers.

```
Nice bar graph of the numbers of missings. The combination graph tells how often combinations of missingness exist. Blue indicate not missing. It shows that there are 3 cases where the only variable missing is age. 1 case where weight and smoker is missing.

c. Another way...matrixplot

```{r}

#windows() #opening matrixplot() in a window allows you to interactively sort by the different variables.
#NOTE! You need to comment out window() if you are going to knit

matrixplot(df[1:3]) #only using the first three columns
```
Red values indicate missingness. COlor of gray boxes indicate some value. White indicates a very low value, with darker gray being a higher value.

If this is all you have done to explore missingness, than you ahve done more than 99% of people. But there are further methods...

Mult. Imputation: take original dataset with missings, create new dataset that fills the missing observations with predicted values that are calculated multiple times....or something.

##Step 5. 
Multiple Imputation using the mice package and functions, and explore the complete cases

```{r}
#reassign as factors

df$sex <- as.factor(df$sex)
df$smoker <- as.factor(df$smoker)

imputed <- mice(df, m = 5) #mice = multiple imputation for change equations, default m is 5
#summary(imputed)

imputed$imp$age

complete(imputed, action = 4)

```
Summary tells you the different types of models that were used to predict values the the complete cases are based on. pmm = predicted means match. Sex does not have a model because there are no missing values. Smoker used logistic regression.

##Step 6. 
Perform regression on each of the complete cases separately using the with() function, and then pool the results using the pool() function

```{r}

lmcc <- with(imputed, lm(sbp ~ weight))
lmcc

#We see 5 different model results. They will be different each time. This randomness helps us explore uncertainty. It shows you the intercept and the coefficient.
 
impute_pooled <- pool(lmcc)
impute_pooled

#The VIM package has the pool() function which allows you to get back to one model that includes this uncertainty that we just created.
```
```{r}

summary(impute_pooled) #Gives us a bunch of info. 
```

