---
title: "Lab 4 Panel Regression"
author: "Brad Anderson"
date: "February 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}

library(tidyverse)
library(Ecdat)
library(plm)
library(car)
library(lmtest)

```

Question: What is the effect of average price of cigarettes on purchases?

2. Simplify the dataset to only retain the variables state, year, packpc, and avgprs.

```{r}

cigs_panel <- Cigarette %>% 
  select(state, year, packpc, avgprs)
```

3. Fixed effects model (within state)

```{r}

cigs_model <- plm(packpc ~ avgprs, 
                  data = cigs_panel,
                  index = c("state", "year"),
                  model = "within")

cigs_model
#summary(cigs_model) ....but we want to use the robust standard error in case of heteroscedasticity

coeftest(cigs_model, vcov = vcovHC(cigs_model, type = "HC1")) #Robust Standard error, Standard error is large, thus a more conservatice approach.
```

Negative relationship between avg price and packs sold per capita. 

Avg price does not significantly predict packs sold of cigarettes.

Null hypothesis (mathematicall) is  the coefficient of average price is 0, which means there is no dependence between the two.


4. Create a Fixed EFfects model for time and entity.

```{r}

et_cigs_model <- plm(packpc ~ avgprs,
                     data = cigs_panel,
                     index = c("state", "year"),
                     model = "within",
                     effect = "twoways") #Instead of just entity fixed effects, we are also inluding the time fixed effect.

et_cigs_model

coeftest(et_cigs_model, vcov =vcovHC(et_cigs_model, type = "HC1")) #vcov is similar to loking at the R squared values of residuals
```

Tells us the same thing, as price increases packs sold goes down.

5. So which of these should I pick? Entity fixed effects or entity/time fixed effects (FE)?

If you only looked at time FE, you would be saying that nothing else changes. This is not realistic in real life. One example could be a VERY controlled lab study. BUt one of the pros of panel data is that you get to look into the ffect of other entities.

```{r}

pFtest(et_cigs_model, cigs_model)


```

This tells us that we should retain the time FE.

6. Least scquares Dummy variable regression (for the states with the top 5 pops in 1995)

This is essentially doing the smae thing as the plm(), but instead creates dummy variables.

```{r}
cigs_panel2 <- Cigarette %>%
  select(state, year, packpc, avgprs) %>% 
  filter(state == "CA" | state == "TX" | state == "NY" | state == "FL" | state == "PA")

```

Least Squares Dummy Variable regression (state FE)

```{r}

lsdv_cigs <- lm(packpc ~ avgprs + state, data = cigs_panel2)
#lsdv_cigs <- lm(packpc ~ avgprs + state + factor*year, data = cigs_panel2) this would include the years as coefficients as well

lsdv_cigs
```

Not the same exact coefficient because we are only using 5 states, but if you ran all the states you should get the same output as the panel regression.

It is nice to have all of the coefficients here for 5 states, but if you had 100s of states it would be cumbersome and panel regression might be a better choice.

```{r}

#a new panel set just for plotting this

cigs_panel_plot <- cigs_panel2 %>% 
  droplevels() #cleans up by removing unobserved levels. If you don't do this R will think there are a bunch of levels from all the states that we filtered out.

```

```{r, fig.height=4, fig.width=4}

scatterplot(cigs_panel_plot$packpc ~ cigs_panel_plot$avgprs | cigs_panel_plot$state,
            xlab = "Average Price", 
            ylab = "Packs per capita",
            smooth = FALSE,
            legend.title = "State") #make a plot of avg cig price vs packs per capit, grouped by state
            

```

