---
title: "Lab 2 - Binary Logistic Regression"
author: "Brad Anderson"
date: "January 27, 2018"
output: html_document
---

message=hide! use it!

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages data, include=FALSE}

library(nnet)
library(leaflet)
library(tidyverse)

badwater_13 <- read_csv("H:/Week 2/badwater_13.csv")

purchase <- read_csv("H:/Week 2/purchase.csv")

```

a. Get our Badwater Ultramarathon 2013 data 

b. Reassigning 'finish and 'sex' as factors 
```{r make_factors}

badwater_13$finish<-as.factor(badwater_13$finish)
badwater_13$sex<-as.factor(badwater_13$sex)

#contrasts(badwater_13$sex) Shows how R assigns numbers to the factors
  #m
#f 0
#m 1

```

```{r badwater model}

bw_model<-glm(finish~sex+age, data=badwater_13, family = "binomial")
summary(bw_model)
# glm is a generalized linear model. it shows a linear relationship between the left had side and right hand side, but what is on the right hand side could be nonlinear (e.g. log odds on left side).


```

Probability of finishing goes down for males compared to females.
Probability of finishing decreases as age goes up.
If you wnat to find the odds, exponentiate the estimated log odds.

e. Make predictions for the model in terms of actual probabilities

Make some data that I will use to make predictions with. 
make range of age. 
one dataframe female 25-75 
second dataframe male 25-75
will take vectors and combine

```{r make predictions}

age_vector<- rep(seq(from=25, to =75), 2)
sex_vector<- c(rep("m",51), rep("f", 51))

new_data<-data.frame(age_vector, sex_vector)
View(new_data) # keep in mind: 
colnames(new_data)<-c("age", "sex")
```

Make somepreductions using the predict function

```{r}

prediction <-predict(bw_model, newdata = new_data, type="response", se.fit = TRUE)
# predict function: choose what you want the calculation to be. if you use type=link, it will give original format on the right hand side of equation (log odds), if you use type=response, turns right hand side of equation into probs. 
#type=link: log odds; type=response --> the actual probabilities
#se.fit=TRUE adds estimate of standard error data

graph_data<-data.frame(new_data, prediction$fit, prediction$se.fit)
View(graph_data)
colnames(graph_data)<-c("age", "sex", "probability", "se")

```

```{r}
# ggplot:needs to know data, what is x and y, and how the graph will look like

ggplot(data=graph_data, aes(x=age,y=probability))+
  geom_line(aes(color=sex))+
  geom_ribbon(aes(ymin=probability-se,  #Press enter after a comma to make a stacked code structure.
                  ymax=probability+se, 
                  fill=sex),alpha=0.3)+
  theme_classic()
```

##2. Multinomial Logistic Regression

a. update to factors 


```{r factors}

purchase$Buy<-as.factor(purchase$Buy)
purchase$EdLevel<-as.factor(purchase$EdLevel)
#look at levels and relevel
purchase$EdLevel<-relevel(purchase$EdLevel, ref="HS") #this resets the reference level, which reorders the way that r see these factors
```

b. Multinomial logistic regression 

```{r multinom}

PurchaseMLR <-multinom(Buy~EdLevel+Income, data=purchase) #How education level and incone influnce Buy. aka Buy as a function of edlevel and income. 
#Because this is multinom() and not glm () we do not need to tell it what kind of regression to run.


```

```{r}

Frame<-data.frame(EdLevel=rep(c("HS","College","Graduate"), 
                              each= 121),
                  Income=rep(seq(0,120),3))
View(Frame)
# NOTE: dont name the dataframe "frame" with lowercase f. View(frame) will not open a dataframe it will open a frame...whatever that is. 
#has to match levels of model

```

d.Make preductions using model 

```{r predict}

Predprob<-data.frame(Frame, predict(PurchaseMLR,newdata=Frame, type="probs"))

```

e.gather() function (like 'melt')

```{r gather}

prob_gather <-Predprob %>% 
  gather(Choice, Probability, Generic:Organic)

```

Graph it!

```{r}

ggplot(prob_gather, aes(x=Income, y=Probability))+
  geom_line(aes(color=Choice))+
  facet_wrap(~EdLevel) #facetwrap: breaks up into different graphs

```

##3. Intro to leaflet

Gets open access data like Openmaps

```{r}

my_map<-leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng=-119.6982, lat=34.4208, popup="Zooming maps")    # tells R you want to open a map

my_map


```
