---
title: "Esm 244 - Lab week 2"
author: "Margaux Sleckman"
date: "January 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


1. Binary Logistic regression 

```{r packages_data}

library(tidyverse)
library(nnet)
library(leaflet)

purchase<-read.csv("C:/Users/Jocelyne/OneDrive/Bren/courses/Winter 2018/ESM 244/244_week2/purchase.csv.csv")
badwater<-read.csv("C:/Users/Jocelyne/OneDrive/Bren/courses/Winter 2018/ESM 244/244_week2/badwater_13.csv.csv")

#View(purchase)
#View(badwater)

```

a. Get our Badwater Ultramarathon 2013 data 

b. Reassigning 'finish and 'sex' as factors 
```{r make_factors}

badwater$finish<-as.factor(badwater$finish)
badwater$sex<-as.factor(badwater$sex)

```

c. Run the binary logistic regression model 

```{r badwater_model}

bw_model<-glm(finish~sex+age, data=badwater, family = "binomial")
summary(bw_model)
# glm shows a linear relationship between the left had side and right hand side, but what is on the left hand side could be nonlinear (e.g. log odds on left side).





```

* note: include = false will not show the data nor the results 
        echo = false will not show the output 
        include=Hide ...
        
Probability of finishing goes down for males compared to females.
Probability of finishing decreases as age goes up.
If you wnat to find the odds, exponentiate the estimated log odds.

e. Make predictions for the model in terms of actual probabilities

Make some data that I will use to make predictions with. 
make range of age. 
one dataframe female 25-75 
second dataframe male 25-75
will tkae vectors and combine 

```{r make_data}

age_vector<- rep(seq(from=25, to =75), 2)
sex_vector<- c(rep("m",51), rep("f", 51))

new_data<-data.frame(age_vector, sex_vector)
View(new_data) # keep in mind: 
colnames(new_data)<-c("age", "sex")


```

Make somepreductions using the predict function

```{r}

prediction <-predict(bw_model, newdata = new_data, type="response", se.fit = TRUE)
# predict function: choose what you want the calculation to be. if you use type=link, it will give original format on the right hand side of equation (log odds), if you use type=response, turns right hand side of equation into probs. 
#type=link: log odds; type=response --> the actual probabilities

graph_data<-data.frame(new_data, prediction$fit, prediction$se.fit)
View(graph_data)
colnames(graph_data)<-c("age", "sex", "probability", "se")

```

g. Graph the probabilities offinishing Badwater

```{r}
# ggplot:needs to know data, what is x and y, and how the graph will look like

ggplot(data=graph_data, aes(x=age,y=probability))+
  geom_line(aes(color=sex))+
  geom_ribbon(aes(ymin=probability-se, ymax=probability+se, fill=sex),alpha=0.3)+
  theme_classic()
```

##2. Multinomial Logistic Regression

a. upadte to factors 


```{r factors}
purchase$Buy<-as.factor(purchase$Buy)
purchase$EdLevel<-as.factor(purchase$EdLevel)
#look at levels and relevel
purchase$EdLevel<-relevel(purchase$EdLevel, ref="HS")

```

b. Multinomial logistic regression 

```{r multinom}

PurchaseMLR <-multinom(Buy~EdLevel+Income, data=purchase)


```

c. Make some new data that we'll have the model make predictions for 

```{r}

Frame<-data.frame(EdLevel=rep(c("HS","College","Graduate"), each= 121), Income=rep(seq(0,120),3))
View(Frame)
# NOTE: dont name the dataframe "frame" with lowercase f. View(frame) will not open a dataframe it will open a frame...whatever that is. 
#has to match levels of model

```

d.Make preductions using model 

```{r predict}

Predprob<-data.frame(Frame, predict(PurchaseMLR,newdata=Frame, type="probs"))

```

e.gather() function (like 'melt')

```{r gather}

prob_gather <-Predprob %>% 
  gather(Choice, Probability, Generic:Organic)

```

Graph it!

```{r}

ggplot(prob_gather, aes(x=Income, y=Probability))+
  geom_line(aes(color=Choice))+
  facet_wrap(~EdLevel)

#facetwrap: breaks up into different graphs
```

##3. Intro to leaflet

```{r}

my_map<-leaflet() %>% 
  addTiles() %>% 
  addMarkers(lng=-119.6982, lat=34.4208, popup="Zooming maps")    # tells R you want to open a map

my_map


```

