---
title: "Lab 6 Time Series & Forecasting"
author: "Brad Anderson"
date: "February 22, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, message=FALSE}

library(tidyverse)
library(forecast)
library(tseries)
```

```{r data}

energy <- read.csv("energy.csv")

class(energy)

```

##3. Convert to 'ts' data
The timeseries package only works with 'ts' class data. The base 'plot' package also understands 'ts' data also, among many others

Our data is measured monthly
```{r }

res_ts <- ts(energy$res_total, frequency = 12, start = c(1973, 1)) #Each year, starting in January 1973

#res_ts

```

```{r}

plot(res_ts)

```

This data is a weird combo of additive and multiplicative data.
Seasonality? YES
Notable Outliers? NO

##4. Decompose the ts data.

decompose() function

```{r}

res_dc <- decompose(res_ts)

plot(res_dc)
```
random = residiuals, looks like white noise, somewhat randomly distributed, the meanas are not changing

##5. Other fast, cool ways to explore your ts data

```{r}

monthplot(res_ts)

```
For each separate month, this shows me how that values has changed over ALL years included in this study. 

Ex. April has not changed a lot, but summer months are showing a drastic increase in energy consumption.
